---
title: Introduction
output: 
   html_document:
    toc: true
    toc_depth: 1
--- 

# Overview
`iviRA` is an R package that runs the . Each model is an individual patient simulation (IPS) that simulates individual patients one at a time. Probabilistic sensitivity analysis (PSA) is used to propagate uncertainty in the input parameters. The simulation is mostly written in `C++` to ensure that the simulation runs in a reasonable amount of time. 

# Installation 
The package is current available on GitHub and can be installed by typing the following commands into the `R` console.

```{r include=TRUE, eval=FALSE}
install.packages("devtools")
library(devtools)
devtools::install_github("InnovationValueInitiative/IVI-RA")
```

Once installed, the package can be loaded into`R`.

```{r }
library("iviRA")
```

# How to run the model
Given a sample of patients and parameters, the IPS is run in three steps. First, the progression of disease is simulated. Second, costs associated with the disease progression are simulated. Third, quality-adjusted life-years (QALYs) are simulated based on the relationship between disease progression and utility. 

<div>
<ul class="pager">
  <li class="next"><a href="02-population.html">Next</a></li>
</ul>
</div>