---
title: The patient population
output: 
   html_document:
    toc: true
    toc_depth: 1
--- 

```{r include=FALSE}
library("iviRA")
```

## Loading the data
Input parameters and data needed to run the simulation come preloaded with the package. To view the loaded data use the command `data(package = "IVI026")`.

## Generating a patient cohort
Before running the IPS, patient characteristics must be specified. The patient dataset must be a matrix containing information on at least patient age, baseline HAQ, gender, and weight. Additionally, if the [Wailoo (2006)](https://www.sheffield.ac.uk/polopoly_fs/1.43356!/file/AHRQ-final-report.pdf) [utility agorithm is used](05-qalys.html), then disease duration and the number of previous DMARDs must be specified as well. 

The patient dataset can be generated using `sample_pats()`. The `type` argument specifies whether the patient cohort is homogeneous or heterogeneous. If the cohort is homogeneous, then the cohort consists of male and female patients that are identical in all respects other than gender; if the cohort is heterogeneous, then all variables vary across patients.

```{r }
pats <- sample_pats(n = 100, type = "heterog")
head(pats)
```

## Input data
To transform the patient dataset into a format needed to run the simulation use the  `input_data` function.

```{r }
input.dat <- input_data(pats)
names(input.dat)
```

<div>
<ul class="pager">
  <li class="previous"><a href="01-intro.html">Previous</a></li>
  <li class="next"><a href="03-parameters.html">Next</a></li>
</ul>
</div>

```{r include=FALSE}
save.image("output.RData")
```