---
title: The patient population
output: 
   html_document:
    toc: true
    toc_depth: 1
--- 

```{r include=FALSE}
library("iviRA")
```

An assessment of a health technology is always conditional on the patient population studied. The first step in an analysis should therefore always be to specifiy the population considered. Below, we consider two ways to generate the patient population: first, the cohort can be randomly sampled using one of the `iviRA` packages built in functions, and, second, you can supply your own patient data. 

## Randomly sampling a patient population 
The simplest way to generate a patient cohort is to use the `sample_pats` function. The variables included in the cohort are age, gender, haq score at baseline, weight, the number of previous dmards, and three measures of disease activity (DAS28, SDAI, and CDAI). The `type` argument specifies whether the patient cohort is homogeneous or heterogeneous. If the cohort is homogeneous, then the cohort consists of male and female patients that are identical in all respects other than gender; if the cohort is heterogeneous, then all variables vary across patients.

We can randomly sample a heterogeneous population with default setting. To examine these settings type `?sample_pats` into the console. 

```{r }
pop <- sample_pats(n = 100, type = "heterog")
head(pop)
```

We can also customize the distribution of patients. For example, suppose we want to study a younger and healthier population. We might then sample patients by customizing arguments in `sample_pats`,

```{r }
pop2 <- sample_pats(n = 100, type = "heterog", age_mean = 45, das28_mean = 4, sdai_mean = 20, 
                    cdai_mean = 18)
head(pop2)
```

## Using your own patient population

<div>
<ul class="pager">
  <li class="previous"><a href="01-intro.html">Previous</a></li>
  <li class="next"><a href="03-parameters.html">Next</a></li>
</ul>
</div>

```{r include=FALSE}
save.image("output.RData")
```