---
title: Sampling the model parameters
output: 
   html_document:
    toc: true
    toc_depth: 1
--- 

```{r include=FALSE}
library("iviRA")
load("output.RData")
```

# Sampling 
PSA is conducted by sampling the parameters from their (joint) probability distribution and running the IPS for each randomly sampled parameter set. The parameters can be sampled using the function `sample_pars`, which generates a probability distribution for all parameters used within the family of models. 

```{r }
parsamp <- sample_pars(n = 100)
names(parsamp)
```

`sample_pars` returns a list containing random draws of the different parameters used in the IPS. For example `parsamp$acr2eular` is an array containing two-way frequency matrices randomly sampled from a Dirichlet distribution that describe the probability of a EULAR response category conditional on ACR response.

```{r }
parsamp$acr2eular[,,1:2]
```

<div>
<ul class="pager">
  <li class="previous"><a href="02-population.html">Previous</a></li>
  <li class="next"><a href="04-ips.html">Next</a></li>
</ul>
</div>

```{r include=FALSE}
save.image("output.RData")
```