---
title: "Generate input data"
output: html_document
---

```{r setup, include=FALSE}
library("iviRA")
library(knitr)
load("output.RData")
```

The IVI-RA model simulates outcomes conditional on parameters and data. Given a patient population, we can generate the required data with the function `get_input_data`

```{r, cache = TRUE }
input.dat <- get_input_data(pop = pop)
names(input.dat)
```

The function returns all patient characteristics needed to run the simulation (HAQ score at baseline, age, gender, DAS28, SDAI, CDAI, weight, and the number of previous DMARDs). It also stores a number of of matrices of explanatory variables. These include:

```{r , echo = FALSE}
library(knitr)
mat <- c("x.mort", "x.acr", "x.haq", "x.das28", "x.ttd.all", "x.ttd.eular", 
         "x.ttd.da", "x.attr")
descr <- c("Variables known at baseline that are used to adjust mortality with a log odds ratio. By default, only the baseline HAQ score is contained in the matrix.", rep("", 7))
parvec <- c("mort.logor", "acr$d", "haq$d", "das28$d", "ttd.all", "ttd.eular",
            "ttd.da", "tx.attr.utility")
table <- cbind(mat, descr, parvec)
colnames(table) <- c("Design matrix", "Description", "Parameters")
kable(table)
```

```{r, cache = TRUE }
head(input.dat$x.mort)
```

The second set of matrices are the design matrices used to simulate time to treatment discontinuation. They include `x.ttd.all`, `x.ttd.da`, and `x.ttd.eular`. The required matrices depnd on the model structures chosen. `x.ttd.all` is required if treatment duration does not depend on disease activity or EULAR respone (i.e., pathway **S1**), `x.ttd.da` is required if treatment duration depends on disease activity (i.e., pathways **S2-S5** ), and `x.ttd.eular` is required if treatment duration depends on EULAR response (i.e., pathway **S6**). This matrix depends on the the model structure chosen since treatment duration depends on the pathway **S1-S6** used for treatment switching. In our selected model structures, `tx_iswitch = acr-eular-switch` or `tx_iswitch = acr-switch`.

```{r, cache = TRUE }
head(input.dat$x.ttd.all)
head(input.dat$x.ttd.eular)
```

<div>
<ul class="pager">
  <li class="previous"><a href="03-structures.html">Previous</a></li>
  <li class="next"><a href="05-parameters.html">Next</a></li>
</ul>
</div>

```{r include=FALSE}
save.image("output.RData")
```