---
title: Simulating Utility and QALYs
output: 
   html_document:
    toc: true
    toc_depth: 1
--- 

```{r include=FALSE}
library("iviRA")
load("output.RData")
```

## Alava (2013) Mixture model
Patient utility as a function of the HAQ score is determined using statistical algorithms that correlate HAQ to utility. The preferred algorithm is the mixture model by [Alava *et al.* 2013](https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/rheumatology/52/5/10.1093_rheumatology_kes400/2/kes400.pdf?Expires=1498633373&Signature=fh0sA4ttg6BOov5xprEAH-BblN8l1DY6lV4LE27ktPAWqQEV2TYL~fY1OuFkvl13L2u8E3xwxaBdyX~zevjuoTdLrxaltGP2S9nbCxcabNDQDrgMDRc0Rk3kIMavbRusV1HEBXxEgXY3Yt7m2-ge5u1MG8okTOgmYR7vQb5oasaA0MxcONrbaIofa5fNi7xrKpIyOo2D50ti1g2aLOJtnzRb5gkI95VJvGUo0pEXglLUem837xCAdKEx~dbMnIecWcGJ8xDsh7xVsIxS72YM5vORh3w26EfXRIPnFzpcWoD3liTkHGXkMhnr4rDtyhsF8PIF3XGFHcQM7MVQHg8gjw__&Key-Pair-Id=APKAIUCZBIA4LVPAVW3Q), which is simulated using `sim_utility_mixture`. Since utility from the mixture model depends on both pain and HAQ, the algorithm simulates `pain` before simulating `utility`.  

```{r, cache = TRUE }
sim.out <- cbind(sim.out, sim_utility_mixture(sim.out, male = input.dat$male, 
                                      pars = c(pain, parsamp$mixture.utility)))
```


## Wailoo (2008) model
The utility mapping algorithm of [Wailoo (2006)](https://www.sheffield.ac.uk/polopoly_fs/1.43356!/file/AHRQ-final-report.pdf) can be used as well. 

```{r, cache = TRUE }
util.wailoo <- sim_utility_wailoo(sim.out, input_data = input.dat, 
                                      pars = parsamp$wailoo.utility)
head(util.wailoo)
```

## Calculate QALYs
QALYs in a given 6-month period are calculated as a function of utility determined by the simulated HAQ score and whether a serious infection occurred (which creates a one-period loss in utility).

```{r }
sim.out$qalys <- sim_qalys(sim.out, sim.out$utility, si_ul = parsamp$si.ul)
head(sim.out[, .(sim, id, month, pain, utility, qalys)])
```

<div>
<ul class="pager">
  <li class="previous"><a href="04-ips.html">Previous</a></li>
  <li class="next"><a href="06-costs.html">Next</a></li>
</ul>
</div>

```{r include=FALSE}
save.image("output.RData")
```