---
title: Performing a cost-effectiveness analysis
output: 
   html_document:
    toc: true
    toc_depth: 1
---

```{r include=FALSE}
library("iviRA")
load("output.RData")
```

The easiest way to analyze output from the `sim_iviRA` function is with the [hesim](http://devinincerti.com/hesim/) package, also developed by IVI. `hesim` helps facilitate computationally intensive simulation modeling, including the analysis of simulation output. It contains a number of functions for summarizing a PSA given general cost-effectiveness output.

To prepare the data so that it can be analyzed with `hesim`, mean outcomes must be calculated for each simulation number and each treatment arm.

```{r }
n.inner <- max(sim.out$sim)
sim.out[, hc_cost := tx_cost + hosp_cost + mgmt_cost + si_cost]
ce <- sim.out[, .(grp = 1,
             qalys = sum(qalys)/n.inner,
                       dqalys = sum(qalys * .03)/n.inner,
                       hc_cost = sum(hc_cost)/n.inner,
                       dhc_cost = sum(hc_cost * .03)/n.inner),
                   by = c("sim", "arm")]
```

The `ce` object can be analyzed in `hesim` using the `pcea` and `pcea_pw` functions, which are designed for "personalized" (i.e., subgroup) cost-effectiveness analysis. Since these functions are designed for subgroup analyses, a group variable must be specified. The current analysis does not specify subgroups so we assume a single large group. Analysis from a health care sector perspective would compare clinical effectiveness as measured by discounted QALYs (`dqalys`) to costs as measured by discounted health care sector costs (`dhc_cost`). An analysis from a societal perspective would add productivity losses to health care sector costs. The 

<div>
<ul class="pager">
  <li class="previous"><a href="05-ips.html">Previous</a></li>
</ul>
</div>
