---
title: Simulating Costs
output: 
   html_document:
    toc: true
    toc_depth: 1
---

```{r include=FALSE}
library("iviRA")
load("output.RData")
```

## Health care sector costs
The costs associated with each arm are simulated using the output from `sim_haq`. To conduct an analysis from a health care perspective, simulate health sector costs using `sim_hc_cost`.

```{r, cache = TRUE }
sim.out <- cbind(sim.out, sim_hc_cost(sim.out, pats[, "weight"], pars = parsamp))
head(sim.out[, .(sim, id, month, infusion_cost, rx_cost, treat_cost, hosp_cost, mgmt_cost, si_cost, hc_cost)])
```

An important component of health care sector costs (`hc_cost`) are the costs associated with treatment (`treat_cost`) with consist of infusion costs (`infusion_cost`) and drug acquisition costs (`rx_cost`). Other health care sector costs are hospital costs (`hosp_cost`), which increase with the HAQ score; general management costs (`mgmt_cost`); and the costs of caused by serious infections (`si_cost`). 

## Productivity losses
To  model currently supports non-health care sector costs due to los wages. Analays performed from a societal perspective should include these productivity losses using the function `sim_prod_loss`.

```{r, cache = TRUE }
sim.out$prod_loss <- sim_prod_loss(sim.out, pl_haq = parsamp$prod.loss)
head(sim.out[, .(sim, id, month, prod_loss)])
```

<div>
<ul class="pager">
  <li class="previous"><a href="05-qalys.html">Previous</a></li>
  <li class="next"><a href="07-cea.html">Next</a></li>
</ul>
</div>

```{r include=FALSE}
save.image("output.RData")
```