---
title: Run the IPS
output: 
   html_document:
    toc: true
    toc_depth: 1
--- 

```{r include=FALSE}
library("iviRA")
load("output.RData")
```


We run the IPS using the function `sim_iviRA`. The main arguments are the selected treatment arms, the input data, the parameter samples, and the model structure. Here, we simulate outcomes for Arm 1 and Arm 2.

```{r }
sim.out1 <- sim_iviRA(arms = arm1, input_data = input.dat, pars = parsamp,
                     model_structures = mod.structs)
sim.out1[, arm := "Arm 1"]
sim.out2 <- sim_iviRA(arms = arm2, input_data = input.dat, pars = parsamp,
                     model_structures = mod.structs)
sim.out2[, arm := "Arm 2"]
sim.out <- rbind(sim.out1, sim.out2)
head(sim.out)
```

`sim_iviRA` returns a data.table which is an enhanced data.frame from the data.table package designed for fast data manipulation. For a given model structure (`model`) and sampled parameter set (`sim`), a sampled patient (`id`) remains on a given therapy until time to treatment discontinuation (`ttd`) becomes less than zero in a given month. Treatment discontinuation is caused by a serious infection (`si = 1`) if the sampled time to serious infection at treatment initiation is less than the sampled time to discontinuation. After discontinuation, a patient switches to the next therapy and `therapy_seq` increments by one. Patient age (`age`) increases in 6-month increments. 

The HAQ score (`haq`) during the initial 6-month period (`therapy_cycle = 1`), disease activity measures (`das28`, `sdai`, and `cdai`), and `ttd` depend on the model structure chosen. For example, in the first model structure, HAQ and time to treatment discontinuation depend on ACR response (`acr`) and in turn, EULAR response (`eular`). The simulation returned NA values for the three disease activity measures because treatment switching is is not a function of disease activity (i.e., **S6** was chosen in the first model structure). The simulation ends when a patient dies (`death = 1`).

Health care sector costs consist of treatment costs (`treat_cost`); hospital costs (`hosp_cost`), which increase with the HAQ score; general management costs (`mgmt_cost`); and the costs of caused by serious infections (`si_cost`). Non-health care sector costs are those due to lost wages. Analyses performed from a societal perspective should include these productivity losses (`prod_loss`). 

Utility (`utility`) depends on whether the Hernandez-Alava mixture model or Wailoo logistic regression equation is used to simulate utility. In both cases, QALYs (`qalys`) are calculated as a function of utility and the estimated utility loss from a serious infection.  

<div>
<ul class="pager">
  <li class="previous"><a href="04-parameters.html">Previous</a></li>
  <li class="next"><a href="06-analysis.html">Next</a></li>
</ul>
</div>

```{r include=FALSE}
save.image("output.RData")
```