---
title: Conducting a cost-effectiveness analysis
output: 
   html_document:
    toc: true
    toc_depth: 1
---

```{r include=FALSE}
library("iviRA")
load("output.RData")
```

## hesim package
The [hesim](http://devinincerti.com/hesim/) package contains a number of functions for conducting cost-effectiveness analyses. To prepare the data so that it can be analyzed with `hesim`, mean outcomes must be calculated for each simulation number and each treatment arm.

```{r }
n.inner <- max(sim.out$sim)
ce <- sim.out[, .(grp = 1,
             qalys = sum(qalys)/n.inner,
                       dqalys = sum(qalys * .03)/n.inner,
                       hc_cost = sum(hc_cost)/n.inner,
                       dhc_cost = sum(hc_cost * .03)/n.inner),
                   by = c("sim", "arm")]
```

The functions in `hesim` are designed for subgroup analyses so we must specify a group variable. The current analysis does not specify subgroups so we assume a single large group. Analysis from a health care sector perspective would compare clinical effectiveness as measured by discounted QALYs (`dqalys`) to costs as measured by discounted health care sector costs (`dhc_cost`). An analysis from a societal perspective would add productivity losses to health care sector costs. 

<div>
<ul class="pager">
  <li class="previous"><a href="06-costs.html">Previous</a></li>
</ul>
</div>
